%% Thesis type: bachelor, master, doctoral
%% Reused by TeXtured metadata even though this document is a report.
\newcommand*{\ThesisType}{report}

%% Document title and metadata
\newcommand*{\ThesisTitle}{Precision Placement in Scaled Dot-Product Attention}
\newcommand*{\ThesisTitlePlaintext}{Precision Placement in Scaled Dot-Product Attention}
\newcommand*{\ThesisTitleFront}{\ThesisTitle}

%% Author
\newcommand*{\ThesisAuthor}{J\'an Kova\v{c}ovsk\'y}
\newcommand*{\ThesisAuthorPlaintext}{Ján Kovačovský}

%% Submission year
\newcommand*{\YearSubmitted}{2026}

%% Basic affiliation/course metadata
\newcommand*{\University}{Charles University}
\newcommand*{\Department}{Faculty of Mathematics and Physics}
\newcommand*{\DeptType}{Faculty}
\newcommand*{\Supervisor}{H. V. Le}
\newcommand*{\SupervisorsDepartment}{Course instructor}
\newcommand*{\StudyProgramme}{NMAG469 -- Mathematical Foundations of Machine Learning}
\newcommand*{\ReportType}{Term Paper Draft}
\newcommand*{\ReportCourse}{Mathematical Foundations of Machine Learning}
\newcommand*{\ReportAffiliation}{Faculty of Mathematics and Physics, Charles University}

%% Abstract
\newcommand*{\Abstract}{%
    We study scaled dot-product attention as a precision-placement problem. Rather than asking whether a single global low-precision dtype is acceptable, we ask which stages of the forward pass can safely use reduced precision: storage of query, key, and value tensors, the logit product, logit storage, rowwise softmax, and the final value product. We prove a local perturbation bound separating logit errors from value errors, and we extend the same viewpoint to quantization by modeling quantized tensors as structured perturbations. Exact-attention experiments show that low-precision storage with high-precision accumulation is comparatively stable, whereas low-precision softmax is the clearest floating-point failure mode. A temperature-scaled attention experiment reveals a finite-size entropy-collapse crossover in which static int8 logit quantization fails through clipping and argmax flips. Finally, an Ising magnetization task gives a preliminary training-time test of whether reduced precision or randomized sketching changes learnability. The sketched-attention baseline is kept as a controlled stress test for the interaction between randomized approximation and finite precision.
}

%% Subject (short description for PDF metadata)
\newcommand*{\Subject}{%
    Precision placement and numerical stability in transformer attention.
}

%% Keywords (about 3-7)
\newcommand*{\Keywords}{%
    mixed-precision arithmetic, attention, transformers, numerical stability, floating-point arithmetic
}
\newcommand*{\KeywordsPlaintext}{%
    mixed-precision arithmetic, attention, transformers, numerical stability, floating-point arithmetic
}
